Reasoning About a Machine with Local Capabilities

### 2018-05-08 toplas/paper.tex
Added some captions and references for the new figures.
The captions need some work, but they illustrate the main idea.

\documentclass[format=acmsmall, review=false, screen=true]{acmart}

@@

\textbf{Returning to the adversary} Perhaps the simplest situation is returning
to the adversary after they invoked our code. In this case, we have received a
return pointer from them, and we just need to jump to it as usual. An obvious
security measure to take care of is properly clearing the non-return-value
registers before we jump (since they may contain data or capabilities that the
adversary should not get access to). Additionally, we may have used the stack
for various purposes (register spilling, storing local state when invoking other
functions etc.), so we also need to clear that data before returning to the
adversary.

However, if we are returning from a function that has itself invoked adversary
code, then clearing the used part of the stack is not enough. The \emph{unused}
part of the stack may also contain data and capabilities, left there by the
adversary, including local capabilities since the stack is write-local. As we
will see later, we rely on the fact that the adversary cannot keep hold of local
capabilities when they pass control to the trusted code and receive control
back. In this case, the adversary could use the unused part of the stack to
store local pointers and load them from there after they get control back. To
prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
the stack that the adversary has had access to, not just the parts that we have
used ourselves. Since we may be talking about a large part of memory, this
requirement is the most problematic aspect of our calling convention for
performance, but see \sectionname~\ref{sec:discussion} for how this might be
mitigated.

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{test}
  \end{subfigure}

  \caption{test}
  \label{fig:ret-adv}
\end{figure}
\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
that we are not using ourselves.

>>>>

\documentclass[format=acmsmall, review=true, screen=true]{acmart}

@@

\textbf{Returning to the adversary} Perhaps the simplest situation is returning
to the adversary after they invoked our code (\figurename~\ref{fig:illu-intro}). In this case, we have received a
return pointer from them, and we just need to jump to it as usual. An obvious
security measure to take care of is properly clearing the non-return-value
registers before we jump (since they may contain data or capabilities that the
adversary should not get access to, \figurename~\ref{fig:illu-priv-data}). Additionally, we may have used the stack
for various purposes (register spilling, storing local state when invoking other
functions etc.), so we also need to clear that data before returning to the
adversary (\figurename~\ref{fig:illu-clear-priv}).

However, if we are returning from a function that has itself invoked adversary
code (\figurename~\ref{fig:illu-call-adv}), then clearing the used part of the stack is not enough. The \emph{unused}
part of the stack may also contain data and capabilities, left there by the
adversary, including local capabilities since the stack is write-local. As we
will see later, we rely on the fact that the adversary cannot keep hold of local
capabilities when they pass control to the trusted code and receive control
back. In this case, the adversary could use the unused part of the stack to
store local pointers and load them from there after they get control back (\figurename~\ref{fig:illu-adv-data}). To
prevent this, we need to clear (i.e.\ overwrite with zeros) the entire part of
the stack that the adversary has had access to, not just the parts that we have
used ourselves (\figurename~\ref{fig:illu-clear-all}). Since we may be talking about a large part of memory, this
requirement is the most problematic aspect of our calling convention for
performance, but see \sectionname~\ref{sec:discussion} for how this might be
mitigated.

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{We have been called by an adversary and are ready to return to the
    adversary.}
  \label{fig:illu-intro}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{If we do not clear our stack when we return to an adversary, then
    they can access our private data and capabilities. This could include old
    stack and program pointers.}
  \label{fig:illu-priv-data}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{Here we made sure to clear the stack before returning to the
    adversary. The adversary no longer has access to our private data and capabilities.}
  \label{fig:illu-clear-priv}
  \end{subfigure}

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,5) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,5)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{In this case, we have called an adversary as part of our program.}
  \label{fig:illu-call-adv}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{If we return to the adversary after the situation in
    \figurename~\ref{fig:illu-call-adv}, the adversary has access to the
    contents of the stack from the call to the adversary. This could include the
    return pointer, we gave to the adversary.}
  \label{fig:illu-adv-data}
  \end{subfigure}
  \begin{subfigure}{0.333\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

      % Stack pointer
      \begin{scope}
        \clip (4.6,2) rectangle (7.6,11.5);
        \capbrace[sp]{(4.6,2)}{(4.6,13)}
      \end{scope}
      \draw node[right=0cm of sp] { sp };
  \end{tikzpicture}
  \caption{Here we made sure to clear all of the stack, including the part we
    gave the adversary access to.}
  \label{fig:illu-clear-all}
  \end{subfigure}

  \caption{In this section, we illustrate the stack in different scenarios
    related to returning to an adversary. }
  \label{fig:ret-adv}
\end{figure}
\textbf{Invoking the adversary} A slightly more complex case is invoking the
adversary. As above, we clear all the non-argument registers, as well as the
part of the stack that we are not using (because, as above, it may contain local
capabilities from previously executed code that the adversary could exploit in
the same way). We leave a copy of the stack pointer in $r_\stk$, but only after
we have used the $\subseg{}{}{}$ instruction to shrink its authority to the part
that we are not using ourselves.
